---
title: "EDS 230 - Assignment 2"
author: "Naomi Moraes & Leilanie Rubinstein"
format: html
editor_options: 
  chunk_output_type: console
---

For this assignment, your goal is to implement a simple model of almond yield anomaly response to climate

- Inputs: daily times series of minimum, maximum daily temperatures and precipitation
- Outputs: maximum, minimum and mean yield anomoly for a input time series

The Lobell et al. 2006 paper will be the source for the transfer function that you will use in your model; specifically look at the equations in table 2.

Here are some ideas to think though. The climate data is going to need to be processed (e.g note that the model uses climate data from a particular month - you have ALL of the climate as daily data). Here are two possible model outlines to follow:

1. Almond_model <- function(clim_var1, clim_var2, parameters) { }
2. Almond_model <- function(clim, parameters) { }

The first example is where the climate variables are separately input into the function - climate data is processed beforehand as part of model set up. The second is where a climate data frame is the input to the function and you extract the useful data from it as part of the model.

You can pick which option you prefer (or try both) If you want a coding challenge, you could create a model that works for any crop.

Run your model for the clim.txt data that is posted on Canvas

`clim.txt` has the following columns: 

These 4 columns tell you when climate observations were made

- day
- month
- year
- wy (water year)

The next 3 columns are the climate observations

- maximum (tmax_c) daily temperature
- minimum (tmin_c) daily temperature
- precipitation (precip) on that day in mm

```{r}
# Load libraries
library(tidyverse)
```

```{r}
# Import climate data
clim <- read_delim(here::here("week2", "data", "clim.txt"))
```

$$
Y = -0.015 T _{n,2} - 0.0046 T _{n,2} ^2 - 0.07 P _1 + 0.0043 P _1 ^2 + 0.28
$$

```{r}
almond_yield <- function(clim, t1 = -0.015, t2 = -0.0046, p1 = -0.07, p2 = 0.0043, c = 0.28) {
  
  # Find minimum Februrary temperature for each year
  min_temp_feb <- clim %>%
    filter(month == 2) %>%
    group_by(year) %>%
    summarise(feb_min_temp = min(tmin_c)) %>%
    ungroup()
  
  # Find January precipitation total for each year
  precip_jan <- clim %>%
    filter(month == 1) %>%
    group_by(year) %>%
    summarise(jan_precip = sum(precip)) %>%
    ungroup()
  
  # Join climate variables
  joined <- left_join(min_temp_feb, precip_jan)
  
  # Calculate almond yield using climate variables
  results <- joined %>%
    mutate(yield = 
             t1*feb_min_temp + 
             t2*feb_min_temp^2 + 
             p1*jan_precip + 
             p2*jan_precip^2 +
             c)
  
  # Calculate yield anomalies
  results <- results %>%
    summarise(min_yield = min(yield), max_yield = max(yield), mean_yield = mean(yield))
}

# Call the almond yield function
almond_yield_anomalies <- almond_yield(clim = clim)

```



















